Sorting by Reversals

### 2020-07-11 paper.tex
update example in paper

\begin{figure}
    \begin{align*}
        \pi = (1\ \mid\ 3\ 4\ \mid\ 2)
    \end{align*}
    \caption{Breakpoints (breakpoints marked by vertical bars)}
    \label{fig:breakpoints}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{tikzpicture}
    	\begin{pgfonlayer}{nodelayer}
    		\node [style=node] (0) at (-3, 0) {0};
    		\node [style=node] (1) at (-1.5, 0) {1};
    		\node [style=node] (2) at (0, 0) {3};
    		\node [style=node] (3) at (1.5, 0) {4};
    		\node [style=node] (4) at (3, 0) {2};
    		\node [style=node] (5) at (4.5, 0) {5};
    	\end{pgfonlayer}
    	\begin{pgfonlayer}{edgelayer}
    		\draw [style=red edge] (1) to (2);
    		\draw [style=red edge] (3) to (4);
    		\draw [style=red edge] (4) to (5);
    		\draw [style=blue edge, bend right] (1) to (4);
    		\draw [style=blue edge, bend left] (2) to (4);
    		\draw [style=blue edge, bend left] (3) to (5);
    	\end{pgfonlayer}
    \end{tikzpicture}
    \caption{Breakpoint graph}
    \label{fig:breakpoint_graph}
\end{figure}

The red edges in the breakpoint graph identify the elements of $\pi$ that have to be split apart and the blue edges identify the elements of $\pi$ that have to be brought together in the process of sorting the permutation $\pi$ (the missing adjacencies). Figure \ref{fig:breakpoint_graph} shows the breakpoint graph of the permutation $(0\ 1\ 3\ 4\ 2\ 5)$.

@@

To give a more visual means of deducing whether an alternating cycle of $G(\pi)$ is oriented, we assign to each red edge $(\pi_i, \pi_{i+1})$ an orientation from $\pi_i$ to $\pi_{i+1}$, orienting red edges of $G(\pi)$ from the endpoint which appears first in $\pi$ to the endpoint which appears second. Then, we call an alternating cycle of $G(\pi)$ \textit{directed} with respect to $\pi$ if it is possible to walk along the whole cycle traversing each red edge in the direction of its orientation. An alternating cycle of $G(\pi)$ is called \textit{undirected} if it is not directed. For example, in figure \ref{fig:breakpoint_graph} the alternating cycle $(\{1,3\},\{2,3\},\{2,5\},\{4,5\},\{2,4\},\{1,2\})$ is directed with respect to $(0\ 1\ 3\ 4\ 2\ 5)$, whereas the alternating cycle $(\{1,3\},\{2,3\},\{2,4\},\{4,5\},\{2,5\},\{1,2\})$ is undirected with respect to $(0\ 1\ 3\ 4\ 2\ 5)$. \citeauthor*{Caprara1997} argues that an alternating cycle of $G(\pi)$ is oriented if and only if it is undirected \cite{Caprara1997}.

@@

\begin{figure}
    \centering
    \begin{tikzpicture}
    	\begin{pgfonlayer}{nodelayer}
    		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
    		\node [style=red node] (1) at (-2, 1) {$\{1,2\}$};
    		\node [style=blue node] (2) at (-1, -1) {$\{4,5\}$};
    		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
    		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
    	\end{pgfonlayer}
    	\begin{pgfonlayer}{edgelayer}
    		\draw (0) to (2);
    		\draw (1) to (2);
    	\end{pgfonlayer}
    \end{tikzpicture}
    \caption{Reversal graph}
    \label{fig:reversal_graph}
\end{figure}

\begin{example}
Consider the alternating cycle decomposition
\begin{align*}
    \mathcal{C} = \{(&\{1,3\},\{2,3\},\{2,4\}, \\
                     &\{4,5\},\{2,5\},\{1,2\})\}
\end{align*}
of the breakpoint graph of $(0\ 1\ 3\ 4\ 2\ 5)$. The corresponding reversal graph is given in figure \ref{fig:reversal_graph}.
\end{example}

@@

\begin{figure}
    \centering
    \begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
    	\begin{pgfonlayer}{nodelayer}
    		\node [style=node] (0) at (-3, 0) {0};
    		\node [style=node] (1) at (-1.5, 0) {1};
    		\node [style=node] (2) at (0, 0) {2};
    		\node [style=node] (3) at (1.5, 0) {4};
    		\node [style=node] (4) at (3, 0) {3};
    		\node [style=node] (5) at (4.5, 0) {5};
    	\end{pgfonlayer}
    	\begin{pgfonlayer}{edgelayer}
    		\draw [style=red edge] (2) to (3);
    		\draw [style=red edge] (4) to (5);
    		\draw [style=blue edge, bend left] (2) to (4);
    		\draw [style=blue edge, bend left] (3) to (5);
    	\end{pgfonlayer}
    \end{tikzpicture}
    \caption{Resulting breakpoint graph}
    \label{fig:breakpoint_graph_change}
\end{figure}

\begin{figure}
    \centering
    \begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
    	\begin{pgfonlayer}{nodelayer}
    		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
    		\node [style=blue node] (1) at (-2, 1) {$\{1,2\}$};
    		\node [style=red node] (2) at (-1, -1) {$\{4,5\}$};
    		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
    		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
    	\end{pgfonlayer}
    	\begin{pgfonlayer}{edgelayer}
    		\draw (0) to (2);
    	\end{pgfonlayer}
    \end{tikzpicture}
    \caption{Resulting reversal graph}
    \label{fig:reversal_graph_change}
\end{figure}

\begin{example}
Applying the reversal represented by vertex $\{1,2\}$ in figure \ref{fig:reversal_graph} results in the breakpoint and reversal graphs shown in figures \ref{fig:breakpoint_graph_change} and \ref{fig:reversal_graph_change} respectively.
\end{example}

\begin{corollary}
A reversal $\rho(u)$ affects only vertices that are in the same connected component as $u$.
\end{corollary}

>>>>

\begin{figure}
    \begin{align*}
        \pi = (7\ \mid\ 5\ 6\ \mid\ 3\ 2\ \mid\ 4\ \mid\ 1)
    \end{align*}
    \caption{Breakpoints (breakpoints marked by vertical bars)}
    \label{fig:breakpoints}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{tikzpicture}
    	\begin{pgfonlayer}{nodelayer}
    		\node [style=node] (0) at (-6, 0) {0};
    		\node [style=node] (1) at (-4.5, 0) {7};
    		\node [style=node] (2) at (-3, 0) {5};
    		\node [style=node] (3) at (-1.5, 0) {6};
    		\node [style=node] (4) at (0, 0) {3};
    		\node [style=node] (5) at (1.5, 0) {2};
    		\node [style=node] (6) at (3, 0) {4};
    		\node [style=node] (7) at (4.5, 0) {1};
    		\node [style=node] (8) at (6, 0) {8};
    	\end{pgfonlayer}
    	\begin{pgfonlayer}{edgelayer}
    		\draw [style=red edge] (0) to (1);
    		\draw [style=red edge] (1) to (2);
    		\draw [style=red edge] (3) to (4);
    		\draw [style=red edge] (5) to (6);
    		\draw [style=red edge] (6) to (7);
    		\draw [style=red edge] (7) to (8);
    		\draw [style=blue edge, bend left] (0) to (7);
    		\draw [style=blue edge, bend left] (1) to (3);
    		\draw [style=blue edge, bend left] (4) to (6);
    		\draw [style=blue edge, bend right] (1) to (8);
    		\draw [style=blue edge, bend right] (2) to (6);
    		\draw [style=blue edge, bend right] (5) to (7);
    	\end{pgfonlayer}
    \end{tikzpicture}
    \caption{Breakpoint graph}
    \label{fig:breakpoint_graph}
\end{figure}

The red edges in the breakpoint graph identify the elements of $\pi$ that have to be split apart and the blue edges identify the elements of $\pi$ that have to be brought together in the process of sorting the permutation $\pi$ (the missing adjacencies). Figure \ref{fig:breakpoint_graph} shows the breakpoint graph of the permutation $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$.

@@

To give a more visual means of deducing whether an alternating cycle of $G(\pi)$ is oriented, we assign to each red edge $(\pi_i, \pi_{i+1})$ an orientation from $\pi_i$ to $\pi_{i+1}$, orienting red edges of $G(\pi)$ from the endpoint which appears first in $\pi$ to the endpoint which appears second. Then, we call an alternating cycle of $G(\pi)$ \textit{directed} with respect to $\pi$ if it is possible to walk along the whole cycle traversing each red edge in the direction of its orientation. An alternating cycle of $G(\pi)$ is called \textit{undirected} if it is not directed. For example, in figure \ref{fig:breakpoint_graph} the alternating cycle $(\{0,7\},\{6,7\},\{3,6\},\{3,4\},\{1,4\},\{0,1\})$ is directed with respect to $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$, whereas the alternating cycle $(\{5,7\},\{4,5\},\{2,4\},\{1,2\},\{1,8\},\{7,8\})$ is undirected with respect to $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$. \citeauthor*{Caprara1997} argues that an alternating cycle of $G(\pi)$ is oriented if and only if it is undirected \cite{Caprara1997}.

@@

\begin{figure}
    \centering
    \begin{tikzpicture}
    	\begin{pgfonlayer}{nodelayer}
    		\node [style=red node] (0) at (-4, 1) {$\{4,5\}$};
    		\node [style=red node] (1) at (-2, 1) {$\{1,2\}$};
    		\node [style=blue node] (2) at (0, 1) {$\{7,8\}$};
    		\node [style=blue node] (3) at (2, 1) {$\{5,6\}$};
    		\node [style=blue node] (4) at (-4, -1) {$\{6,7\}$};
    		\node [style=blue node] (5) at (-2, -1) {$\{3,4\}$};
    		\node [style=blue node] (6) at (0, -1) {$\{0,1\}$};
    		\node [style=blue node] (7) at (2, -1) {$\{2,3\}$};
    	\end{pgfonlayer}
    	\begin{pgfonlayer}{edgelayer}
    		\draw (0) to (4);
    		\draw (0) to (1);
    		\draw (0) to (5);
    		\draw (1) to (5);
    		\draw (1) to (6);
    		\draw (2) to (4);
    		\draw (2) to (6);
    	\end{pgfonlayer}
    \end{tikzpicture}
    \caption{Reversal graph}
    \label{fig:reversal_graph}
\end{figure}

\begin{example}
Consider the alternating cycle decomposition
\begin{align*}
    \mathcal{C} = \{&(\{0,7\},\{6,7\},\{3,6\},\{3,4\},\{1,4\},\{0,1\}), \\
                    &(\{5,7\},\{4,5\},\{2,4\},\{1,2\},\{1,8\},\{7,8\})\}
\end{align*}
of the breakpoint graph of $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$. The corresponding reversal graph is given in figure \ref{fig:reversal_graph}.
\end{example}

@@

\begin{figure}
    \centering
    \begin{tikzpicture}
    	\begin{pgfonlayer}{nodelayer}
    		\node [style=node] (0) at (-6, 0) {0};
    		\node [style=node] (1) at (-4.5, 0) {7};
    		\node [style=node] (2) at (-3, 0) {2};
    		\node [style=node] (3) at (-1.5, 0) {3};
    		\node [style=node] (4) at (0, 0) {6};
    		\node [style=node] (5) at (1.5, 0) {5};
    		\node [style=node] (6) at (3, 0) {4};
    		\node [style=node] (7) at (4.5, 0) {1};
    		\node [style=node] (8) at (6, 0) {8};
    	\end{pgfonlayer}
    	\begin{pgfonlayer}{edgelayer}
    		\draw [style=red edge] (0) to (1);
    		\draw [style=red edge] (1) to (2);
    		\draw [style=red edge] (3) to (4);
    		\draw [style=red edge] (6) to (7);
    		\draw [style=red edge] (7) to (8);
    		\draw [style=blue edge, bend left] (0) to (7);
    		\draw [style=blue edge, bend left] (1) to (4);
    		\draw [style=blue edge, bend left] (3) to (6);
    		\draw [style=blue edge, bend right] (1) to (8);
    		\draw [style=blue edge, bend right] (2) to (7);
    	\end{pgfonlayer}
    \end{tikzpicture}
    \caption{Resulting breakpoint graph}
    \label{fig:breakpoint_graph_change}
\end{figure}

\begin{figure}
    \centering
    \begin{tikzpicture}
    	\begin{pgfonlayer}{nodelayer}
    		\node [style=blue node] (0) at (-4, 1) {$\{4,5\}$};
    		\node [style=blue node] (1) at (-2, 1) {$\{1,2\}$};
    		\node [style=blue node] (2) at (0, 1) {$\{7,8\}$};
    		\node [style=blue node] (3) at (2, 1) {$\{5,6\}$};
    		\node [style=red node] (4) at (-4, -1) {$\{6,7\}$};
    		\node [style=red node] (5) at (-2, -1) {$\{3,4\}$};
    		\node [style=blue node] (6) at (0, -1) {$\{0,1\}$};
    		\node [style=blue node] (7) at (2, -1) {$\{2,3\}$};
    	\end{pgfonlayer}
    	\begin{pgfonlayer}{edgelayer}
    		\draw (1) to (4);
    		\draw (1) to (6);
    		\draw (2) to (4);
    		\draw (2) to (6);
    		\draw (4) to (5);
    	\end{pgfonlayer}
    \end{tikzpicture}
    \caption{Resulting reversal graph}
    \label{fig:reversal_graph_change}
\end{figure}

\begin{example}
Applying the reversal represented by vertex $\{4,5\}$ in figure \ref{fig:reversal_graph} results in the breakpoint and reversal graphs shown in figures \ref{fig:breakpoint_graph_change} and \ref{fig:reversal_graph_change} respectively.
\end{example}

\begin{corollary}
A reversal $\rho(u)$ affects only vertices that are in the same connected component as $u$.
\end{corollary}\padding
